---
title: "Assignment 5"
author: "Saima Rahmanzai"
date: "4/29/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


## Set the working directory to the root of your DSC 520 directory
##setwd("/home/jdoe/Workspaces/dsc520")

options(max.print=10000)

setwd("C:\\Users\\Saima\\Class_Assignments\\dsc520")


## Load the `data/r4ds/heights.csv` to
heights_df <- read.csv("data/r4ds/heights.csv")
head(heights_df)

## Using `cor()` compute correlation coefficients for
## height vs. earn

cor(heights_df$height,heights_df$earn)

### age vs. earn

cor(heights_df$age,heights_df$earn)

### ed vs. earn
cor(heights_df$ed,heights_df$earn)

## Spurious correlation
## The following is data on US spending on science, space, and technology in millions of today's dollars
## and Suicides by hanging strangulation and suffocation for the years 1999 to 2009
## Compute the correlation between these variables
tech_spending <- c(18079, 18594, 19753, 20734, 20831, 23029, 23597, 23584, 25525, 27731, 29449)
suicides <- c(5427, 5688, 6198, 6462, 6635, 7336, 7248, 7491, 8161, 8578, 9000)

##cor() computes the correlation coefficient
##cor.test() test for association/correlation between paired samples. It returns both the correlation coefficient 
##and the significance level(or p-value) of the correlation
##use = "complete.obs" removes missing values

cor(tech_spending, suicides, method = c("pearson"), use = "complete.obs")

cor.test(tech_spending, suicides, method=c("pearson"), use = "complete.obs")

## The above can be visualized in a graph as well using the below commands

library("ggpubr")
ggqqplot(tech_spending, ylab = "US Spending on Science, space, and technology")
ggqqplot(suicides, ylab = "Suicides by hanging, strangulation, and suffocation")
##above shows both data come from normal distribution so the pearson correlation works.

shapiro.test(tech_spending)
##This test shows if p value is greater than the significance level of .05 implying that the distribution of the data
## is not significantly different from normal distribution.  In other words, we can assume normality as the p value is 0.6285
## for tech_spending.


shapiro.test(suicides)
##the p value for suicides is 0.8674 which is also greater than the significance level of .05 implying that the distribution
##of data is not significantly different from normal distribution.




theDF <- data.frame(tech_spending, suicides)

theDF

library(ggplot2)
g<-ggplot(theDF, aes(x=suicides,y=tech_spending)) + geom_point()+ geom_smooth(formula = y~x, method='lm')
g + xlab("Suicides by hanging, strangulation, and suffocation") +
    ylab("US Spending on Science, Space, and Technology") +
    ggtitle("US Spending on Science, Space & Technology Vs. Suicides Correlation")


theDF <- data.frame(tech_spending, suicides)

theDF

library("ggpubr")


ggscatter(theDF, x = "suicides", y = "tech_spending", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Suicides", ylab = "US Spending",
          geom_smooth(formula = y~x))







##The relationship is linear. In the situation where the scatter plots show curved patterns, we are dealing with nonlinear association between the two variables.
## We ran Use Shapiro-Wilk normality test and the ggqplot test for normality.  Appears to be a normal distribution.
```

## Set the working directory to the root of your DSC 520 directory

```{r results = 'asis', echo=TRUE} 
options(max.print=10000)

setwd("C:\\Users\\Saima\\Class_Assignments\\dsc520")
```


## Load the data/r4ds/heights.csv

```{r results = 'asis', echo=TRUE} 

heights_df <- read.csv("data/r4ds/heights.csv")
head(heights_df)
```



## Using cor() compute correlation coefficients for
## height vs. earn

```{r results = 'asis', echo=TRUE} 
cor(heights_df$height,heights_df$earn)
```

### age vs. earn

```{r results = 'asis', echo=TRUE}
cor(heights_df$age,heights_df$earn)
```


### ed vs. earn

```{r results = 'asis', echo=TRUE}
cor(heights_df$ed,heights_df$earn)
```


## Spurious correlation
## The following is data on US spending on science, space, and technology in millions of today's dollars
## and Suicides by hanging strangulation and suffocation for the years 1999 to 2009
## Compute the correlation between these variables
## tech_spending <- c(18079, 18594, 19753, 20734, 20831, 23029, 23597, 23584, 25525, 27731, 29449)
## suicides <- c(5427, 5688, 6198, 6462, 6635, 7336, 7248, 7491, 8161, 8578, 9000)

```{r results = 'asis', echo=TRUE}

tech_spending <- c(18079, 18594, 19753, 20734, 20831, 23029, 23597, 23584, 25525, 27731, 29449)
suicides <- c(5427, 5688, 6198, 6462, 6635, 7336, 7248, 7491, 8161, 8578, 9000)

cor(tech_spending, suicides, method = c("pearson"), use = "complete.obs")

cor.test(tech_spending, suicides, method=c("pearson"), use = "complete.obs")

```


## The above can be visualized in a graph as well using the below commands

```{r results = 'asis', echo=TRUE}
library("ggpubr")
ggqqplot(tech_spending, ylab = "US Spending on Science, space, and technology")
ggqqplot(suicides, ylab = "Suicides by hanging, strangulation, and suffocation")

```

## Above shows both data come from normal distribution so the pearson correlation works.


```{r results = 'asis', echo=TRUE}
shapiro.test(tech_spending)
```
## This test shows if p value is greater than the significance level of .05 implying that the distribution of the data is not significantly different from normal distribution.  In other words, we can assume normality as the p value is 0.6285 tech_spending.

```{r results = 'asis', echo=TRUE}
shapiro.test(suicides)
```

## The p value for suicides is 0.8674 which is also greater than the significance level of .05 implying that the distribution of data is not significantly different from normal distribution.


## Including Plots

The scatterplots I generated to show relationships between two variables are provided below:


## The relationship is linear. In the situation where the scatter plots show curved patterns, we are dealing with nonlinear association between the two variables.We ran Shapiro-Wilk normality test abpve and the ggqplot test below for normality.  Appears to be a normal distribution and linear.




